---
title: "Customer Database & Sales Team Analysis"
subtitle: "Record-Keeping & Account Management Report"
author: "Data Analysis Team"
date: "June 05, 2026"
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    theme: flatly
    embed-resources: true
    fig-align: center
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(knitr)

shola_clean        <- readRDS("shola_clean.rds")
ian_detail         <- readRDS("ian_detail.rds")
peter_detail       <- readRDS("peter_detail.rds")
city_summary       <- readRDS("city_summary.rds")
rep_region         <- readRDS("rep_region.rds")
rep_summary_export <- readRDS("rep_summary_export.rds")
```

## Executive Summary

This report presents the findings from a full audit of the company's customer
database. The analysis covers **2,273 verified customer records** across
**31 active sales representatives** and **20+ regions** nationwide.

::: {.callout-important}
## Three Actions Requiring Immediate CEO Attention

1. **441 customers (19.4% of the base)** are managed by just two reps — Ian and
   Peter — with no backup or succession plan in place.
2. **341 customers** are assigned to a placeholder entry called *"Customer Office"*
   and have no named sales representative.
3. **Address and contact data gaps** exist across the database — 9.3% of customers
   have no street address and 4.7% have no phone number.
:::

---

## Dataset Overview

### Cleaning Summary

The raw database required significant cleaning before analysis. The table below
summarises what was found and what was done.

```{r}
#| label: tbl-cleaning
#| tbl-cap: "Data Cleaning Steps Applied to Raw Database"
tibble::tribble(
  ~Step,  ~Issue,                                          ~Action,
  "1",  "Metadata rows at top of file",                  "Removed 5 header/metadata rows",
  "2",  "Incorrect column names",                        "Set correct names from row 5",
  "3",  "Mixed-case text in all columns",                "Trimmed whitespace; applied title case to rep names",
  "4",  "3 near-empty columns (100% / 99.6% missing)",   "Dropped Address Line 2, Address Line 3, Post Code",
  "5",  "Fax Number 70% missing",                        "Dropped Fax Number column",
  "6",  "6 junk rows (Hebrew text, notes, no customer ID)", "Removed non-customer rows",
  "7",  "3 suspect sales rep entries (dummy IDs, Hebrew placeholders)", "Removed כללי, Oded, DR MUSILIU AGBOOLA",
  "8",  "Remaining rows with any missing value",          "Removed; 2,273 complete records retained"
) |>
  kable(col.names = c("Step", "Issue Found", "Action Taken"))
```

### Final Dataset Dimensions

```{r}
#| label: tbl-dims
#| tbl-cap: "Final Clean Dataset at a Glance"
tibble::tribble(
  ~Metric,               ~Value,
  "Total customer records",   "2,273",
  "Columns retained",         "7",
  "Active sales reps",        "31",
  "Regions covered",          "20+",
  "Missing values remaining", "0"
) |>
  kable(col.names = c("Metric", "Value"))
```

---

## Sales Team Analysis

### Customers Per Sales Representative

@fig-rep-count shows the number of customers each sales representative currently
manages, ranked from highest to lowest.

```{r}
#| label: fig-rep-count
#| fig-cap: "Number of customers managed per sales representative (all reps)"
#| fig-height: 8
shola_clean |>
  count(`Sales Rep Name`, sort = TRUE) |>
  ggplot(aes(x = n, y = reorder(`Sales Rep Name`, n))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Customers", y = NULL)
```

::: {.callout-note}
## Key Observations

- **"Customer Office"** (341 customers) is a placeholder — not a real rep.
  These accounts are effectively unmanaged.
- **Ian (232) and Peter (209)** carry the heaviest legitimate workloads,
  each managing roughly twice the number of customers of any other rep.
- **5 reps manage only 1 customer each** — these should be verified as active.
:::

### Workload Distribution — Real Reps Only

```{r}
#| label: fig-rep-count-clean
#| fig-cap: "Customer count per rep — excluding 'Customer Office' placeholder"
#| fig-height: 8
shola_clean |>
  filter(`Sales Rep Name` != "Customer Office") |>
  count(`Sales Rep Name`, sort = TRUE) |>
  ggplot(aes(x = n, y = reorder(`Sales Rep Name`, n))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Customers", y = NULL)
```

---

## Geographic Distribution

### Where Are Our Customers?

@fig-geo shows how customers are distributed across cities and states nationwide,
after consolidating inconsistent location entries from the raw database.

```{r}
#| label: fig-geo
#| fig-cap: "Customer distribution by location — top 20 cities / states"
#| fig-height: 7
ggplot(city_summary, aes(x = n, y = reorder(location, n))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Customers", y = NULL)
```

::: {.callout-note}
## Geographic Highlights

- **Ibadan / Oyo** is the dominant market with 711 customers — nearly 3× the
  next largest location.
- **Ogun State (236), Lagos (197), and Kaduna (158)** form the next tier.
- Southern and middle-belt states hold 70–95 customers each, suggesting broad
  national reach but thin coverage outside the South-West.
:::

---

## Regional Coverage by Sales Rep

@fig-heatmap maps which reps are active in which regions, using the top 10
locations and top 15 reps. Darker cells indicate more customers.

```{r}
#| label: fig-heatmap
#| fig-cap: "Sales rep coverage heatmap — top 10 regions × top 15 reps"
#| fig-height: 7
#| fig-width: 12
ggplot(rep_region,
       aes(x = `Sales Rep Name`, y = reorder(location, n, sum), fill = n)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = ifelse(n > 0, n, "")), size = 2.8, color = "white") +
  scale_fill_gradient(low = "#d4e6f1", high = "#1a5276", name = "Customers") +
  labs(x = NULL, y = NULL) +
  theme(
    axis.text.x  = element_text(angle = 40, hjust = 1, size = 8),
    axis.text.y  = element_text(size = 9)
  )
```

::: {.callout-warning}
## Territory Overlap Concerns

- Most regions have **multiple reps sharing the same territory** with no clear
  ownership — creating confusion and duplicated effort.
- **Peter** is the only rep active across 6+ regions, creating a single point
  of failure for the South-West.
- **Ian** dominates the entire South-East corridor across 31 regions.
:::

---

## Key Account Management Risks {#sec-risks}

### Risk Overview

```{r}
#| label: tbl-risk
#| tbl-cap: "Ian vs Peter — Side-by-Side Risk Comparison"
tibble::tribble(
  ~Metric,                          ~Ian,                        ~Peter,
  "Total customers managed",        "232",                       "209",
  "Regions covered",                "31",                        "22",
  "Regions with only 1 customer",   "13",                        "13",
  "Largest single region",          "Enugu — 51 (22%)",          "Ibadan — 115 (55%)",
  "Portfolio concentration",        "Low (evenly spread)",       "High (Ibadan-heavy)",
  "Geographic coherence",           "Strong (South-East corridor)", "Weak (scattered)",
  "Succession difficulty",          "Very High",                 "High"
) |>
  kable(col.names = c("Metric", "Ian", "Peter"))
```

### Ian — Regional Spread

```{r}
#| label: fig-ian
#| fig-cap: "Ian — 232 customers across 31 regions"
#| fig-height: 9
ggplot(ian_detail, aes(x = n, y = reorder(location, n))) +
  geom_col() +
  geom_text(aes(label = paste0(n, "  (", Pct, "%)")), hjust = -0.1, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25))) +
  labs(
    subtitle = "232 customers | 31 regions | South-East corridor specialist",
    x = "Number of Customers", y = NULL
  )
```

::: {.callout-important}
## Ian Risk Summary

Ian is the business's largest rep by customer count. His portfolio spans
**31 regions** — the widest of any rep — with **no dominant anchor region**
(largest region: Enugu at only 22%). If Ian left, accounts would need to be
redistributed across 31 different regions simultaneously, making him the
**highest succession risk** in the business.
:::

### Peter — Regional Spread

```{r}
#| label: fig-peter
#| fig-cap: "Peter — 209 customers across 22 regions"
#| fig-height: 9
ggplot(peter_detail, aes(x = n, y = reorder(location, n))) +
  geom_col() +
  geom_text(aes(label = paste0(n, "  (", Pct, "%)")), hjust = -0.1, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25))) +
  labs(
    subtitle = "209 customers | 22 regions | 55% concentrated in Ibadan/Oyo",
    x = "Number of Customers", y = NULL
  )
```

::: {.callout-important}
## Peter Risk Summary

Peter manages **115 customers in Ibadan/Oyo alone** — the business's single
largest market (55% of his portfolio). He is also the de facto owner of this
territory. Losing Peter would mean the largest customer hub in the database
loses its primary relationship manager overnight. His 13 single-customer
outlier regions (including Ghana, Kaduna, Benue) are geographically incoherent
and likely receiving minimal attention.
:::

---

## Priority Action Plan for the CEO

```{r}
#| label: tbl-actions
#| tbl-cap: "Recommended Actions — Prioritised for CEO"
tibble::tribble(
  ~Priority, ~Action,                                                                   ~Owner,           ~Timeline,
  "1",       "Identify a South-East deputy for Ian across his core 5 regions",          "CEO / HR",       "Immediate",
  "2",       "Identify a named backup rep for Peter's 115 Ibadan/Oyo accounts",         "CEO / HR",       "Immediate",
  "3",       "Assign the 341 'Customer Office' accounts to named sales reps",           "Sales Manager",  "Within 1 week",
  "4",       "Redistribute Ian and Peter's 26 combined single-customer outlier regions","Sales Manager",  "Within 2 weeks",
  "5",       "Cap both Ian and Peter's portfolios at 150 accounts",                     "Sales Manager",  "Within 1 month",
  "6",       "Define formal territory maps for all 31 reps",                            "Sales Manager",  "Within 1 month",
  "7",       "Fill missing street address and phone number gaps (9.3% and 4.7%)",       "Admin Team",     "Within 1 month"
) |>
  kable(col.names = c("Priority", "Action", "Owner", "Suggested Timeline"))
```

---

## Appendix — Full Sales Rep Summary

```{r}
#| label: tbl-reps-full
#| tbl-cap: "All Sales Representatives — Customer Count and Notes"
rep_summary_export |>
  kable(col.names = c("Rank", "Rep No.", "Sales Rep Name", "Customer Count", "Note"))
```

---

*Report prepared by the Data Analysis Team | Data current as of 05 June 2026*
*Source: shola_clean.xlsx — 2,273 verified customer records*
