---
title: "Customer & Sales Rep Analysis"
subtitle: "Agro Bar Magen Nigeria LTD"
author: "Akinbami Oluwashola"
date: "June 2026"
format:
  beamer:
    theme: Madrid
    colortheme: dolphin
    fonttheme: structurebold
    aspectratio: 169
    fontsize: 10pt
    navigation: horizontal
    section-titles: true
    toc: false
    header-includes: |
      \setbeamertemplate{footline}[frame number]
      \usepackage{booktabs}
      \usepackage{longtable}
      \definecolor{abmgreen}{RGB}{34,139,34}
      \definecolor{abmblue}{RGB}{0,70,140}
      \setbeamercolor{title}{fg=abmblue}
      \setbeamercolor{frametitle}{fg=abmblue}
execute:
  echo: false
  warning: false
  message: false
  fig-width: 8
  fig-height: 4.2
---

```{r}
#| label: load-packages
library(tidyverse)
library(readxl)
library(knitr)
library(scales)
```

```{r}
#| label: load-data
shola_clean <- readRDS("shola_clean.rds")
rep_summary <- readRDS("rep_summary_export.rds")
city_summary <- readRDS("city_summary.rds")
ian_detail <- readRDS("ian_detail.rds")
peter_detail <- readRDS("peter_detail.rds")
rep_region <- readRDS("rep_region.rds")
```

```{r}
#| label: prepare-data
total_customers <- nrow(shola_clean)
total_reps <- n_distinct(shola_clean$`Sales Rep Name`, na.rm = TRUE)
total_locations <- n_distinct(city_summary$location)

# Overlap data
overlap <- ian_detail |>
  inner_join(peter_detail, by = "location", suffix = c("_ian", "_peter")) |>
  rename(ian_n = n_ian, ian_pct = Pct_ian, peter_n = n_peter, peter_pct = Pct_peter) |>
  mutate(combined_customers = ian_n + peter_n) |>
  arrange(desc(combined_customers))

overlap_ian <- sum(overlap$ian_n)
overlap_peter <- sum(overlap$peter_n)
```

# Executive summary

## Executive summary

\begin{center}
\large\textbf{Key numbers at a glance}
\end{center}

\vspace{0.3cm}

| Metric | Value |
|:-------|------:|
| Total customers on record | `{r} format(total_customers, big.mark = ",")` |
| Active sales representatives | `{r} total_reps` |
| Geographic locations covered | `{r} total_locations`+ |

\vspace{0.3cm}

**Top 3 reps by customer count:**

1. Customer Office — 341 customers (administrative allocation)
2. **Ian** — 232 customers across 31 locations (widest geographic spread)
3. **Peter** — 209 customers across 22 locations (heavily concentrated)

## Key findings

- **Geographic concentration:** Over 31% of all customers are in Ibadan/Oyo alone.
- **Ian** has the widest territorial spread but moderate concentration — his top 3 locations hold \~54% of his portfolio.
- **Peter** is heavily concentrated — Ibadan/Oyo alone holds **55%** of his customers.
- Ian and Peter share **7 locations**, but actual competitive overlap is minimal — in each shared location, one rep dominates.
- **Data quality:** 9.5% of customer records lack a street address; 4.9% are missing phone numbers.

# Data overview

## Data source & scope

**Source:** Company customer database export (Excel)

**Scope:** All customers mapped to their assigned sales representatives.

\vspace{0.3cm}

**Raw data:** 2,604 rows × 11 columns (all character type)

**After cleaning:** `{r} format(total_customers, big.mark = ",")` rows × 7 columns

\vspace{0.3cm}

**Cleaning steps applied:**

- Removed header/metadata rows (company name, timestamp, report title)
- Dropped near-empty columns (Address Line 2, State, Fax Number — \~100% missing)
- Standardised sales rep names to consistent casing
- Grouped cities/counties into cleaner location labels
- Converted Customer Number and Sales Rep Number to numeric

## Data quality summary

| Column | Missing count | Missing % |
|:-------|-----:|-----:|
| Street Address | 247 | 9.5% |
| Phone Number | 127 | 4.9% |
| City/County | 126 | 4.8% |
| Sales Rep Number | 11 | 0.4% |
| Customer Number | 6 | 0.2% |
| Customer Name | 0 | 0.0% |

\vspace{0.3cm}

**Recommendation:** Prioritise filling in missing street addresses and phone numbers to improve customer reachability and delivery accuracy.

# Customer distribution by location

## Top 20 customer locations

```{r}
#| label: fig-location-dist
#| fig-cap: "Customer count by location (Top 20)"
#| fig-height: 4.5

city_summary |>
  mutate(location = fct_reorder(location, n)) |>
  ggplot(aes(x = n, y = location)) +
  geom_col(fill = "#00468B") +
  geom_text(aes(label = n), hjust = -0.1, size = 3) +
  labs(x = "Number of Customers", y = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank())
```

## Location analysis — Key takeaways

- **Ibadan/Oyo dominates** with 711 customers (31% of all customers) — both an opportunity and a concentration risk.
- **Top 5 locations** (Ibadan/Oyo, Ogun, Lagos, Kaduna, Ilorin/Kwara) account for roughly **62%** of all customers.
- Southern Nigeria is heavily represented; northern presence is limited mainly to Kaduna and Kano.

\vspace{0.3cm}

**Implications for leadership:**

- Revenue is geographically concentrated — disruptions in Ibadan/Oyo would disproportionately affect the business.
- There may be untapped growth opportunities in under-represented regions (e.g., Abuja/FCT, northern states).

# Sales rep performance

## Sales reps ranked by customer count (Top 15)

```{r}
#| label: fig-rep-ranking
#| fig-cap: "Top 15 sales reps by customer count"
#| fig-height: 4.5

rep_summary |>
  head(15) |>
  mutate(`Sales Rep Name` = fct_reorder(`Sales Rep Name`, `Customer Count`)) |>
  ggplot(aes(x = `Customer Count`, y = `Sales Rep Name`)) +
  geom_col(fill = "#00468B") +
  geom_text(aes(label = `Customer Count`), hjust = -0.1, size = 3) +
  labs(x = "Number of Customers", y = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank())
```

## Sales rep performance — Key takeaways

- **Customer Office** (341 customers) is an administrative allocation, not a single sales rep.
- **Ian (232) and Peter (209)** are the top two individual performers by customer count.
- There is a steep drop-off after the top 5 reps: the **top 5 account for a disproportionate share** of the customer base.
- **11 reps have fewer than 30 customers each** — these may be newer hires, part-time, or underperforming.

\vspace{0.3cm}

**Recommendations:**

- Investigate why Customer Office holds so many accounts — consider redistributing to named reps.
- Review workload balance: top reps may be stretched too thin while others are underutilised.

# Deep dive: Ian

## Ian's customer distribution by location

```{r}
#| label: fig-ian-detail
#| fig-cap: "Ian's 232 customers across 31 locations"
#| fig-height: 4.5

ian_detail |>
  slice_max(n, n = 15) |>
  mutate(location = fct_reorder(location, n)) |>
  ggplot(aes(x = n, y = location)) +
  geom_col(fill = "#2E8B57") +
  geom_text(aes(label = paste0(n, " (", Pct, "%)")), hjust = -0.05, size = 3) +
  labs(x = "Number of Customers", y = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank())
```

## Ian — Analysis & risk assessment

**Strengths:**

- Widest geographic coverage of any rep — **31 locations**.
- No single location exceeds 22% of his portfolio.
- Strong presence across the South-East and South-South (Enugu, Delta, Rivers, Edo).

\vspace{0.2cm}

**Risks:**

- His top 3 locations (Enugu, Delta, Rivers/PH) hold \~54% of his customers — moderate concentration.
- Managing 31 locations may strain his ability to service all customers effectively.
- If Ian were to leave, the company would lose coverage in numerous locations simultaneously.

\vspace{0.2cm}

**Recommendations:**

- Develop a succession plan for Ian's territory.
- Consider assigning a junior rep to shadow Ian in his top locations.
- Monitor customer satisfaction in Ian's lower-volume locations.

# Deep dive: Peter

## Peter's customer distribution by location

```{r}
#| label: fig-peter-detail
#| fig-cap: "Peter's 209 customers across 22 locations"
#| fig-height: 4.5

peter_detail |>
  slice_max(n, n = 15) |>
  mutate(location = fct_reorder(location, n)) |>
  ggplot(aes(x = n, y = location)) +
  geom_col(fill = "#B8860B") +
  geom_text(aes(label = paste0(n, " (", Pct, "%)")), hjust = -0.05, size = 3) +
  labs(x = "Number of Customers", y = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank())
```

## Peter — Analysis & risk assessment

**Strengths:**

- Second-highest individual customer count (209).
- Dominant presence in the company's largest market — Ibadan/Oyo.

\vspace{0.2cm}

**Critical risk — Geographic concentration:**

- **55% of Peter's customers** are in Ibadan/Oyo alone.
- His top 2 locations (Ibadan/Oyo + Ogun) account for **72%** of his portfolio.
- If Peter loses access to Ibadan/Oyo (e.g., he relocates, becomes ill, or resigns), 115 customers are immediately at risk.

\vspace{0.2cm}

**Recommendations:**

- **Urgent:** Assign a backup rep for Ibadan/Oyo to reduce single-person dependency.
- Gradually redistribute some Ibadan/Oyo accounts to reduce Peter's concentration.
- Encourage Peter to develop his presence in less-saturated markets.

# Ian vs Peter comparison

## Side-by-side: Top 10 locations

```{r}
#| label: fig-ian-peter-compare
#| fig-cap: "Top 10 locations for Ian and Peter (side by side)"
#| fig-height: 4.5

ian_top <- ian_detail |> slice_max(n, n = 10) |> mutate(rep = "Ian")
peter_top <- peter_detail |> slice_max(n, n = 10) |> mutate(rep = "Peter")

bind_rows(ian_top, peter_top) |>
  mutate(location = fct_reorder(location, n, .fun = sum)) |>
  ggplot(aes(x = n, y = location, fill = rep)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Ian" = "#2E8B57", "Peter" = "#B8860B")) +
  labs(x = "Number of Customers", y = NULL, fill = "Sales Rep") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        legend.position = "top")
```

## Comparison summary

| Metric | Ian | Peter |
|:-------|----:|------:|
| Total customers | 232 | 209 |
| Locations covered | 31 | 22 |
| Top location share | 22% (Enugu) | 55% (Ibadan/Oyo) |
| Top 3 locations share | \~54% | \~81% |

\vspace{0.3cm}

**Key contrast:** Ian is a *spread* rep — wide coverage, moderate concentration. Peter is a *depth* rep — deep penetration in fewer markets. Both strategies have merit, but Peter's concentration poses higher business continuity risk.

## Overlap analysis

Ian and Peter share **7 locations** out of 46 unique locations combined.

```{r}
#| label: tbl-overlap
#| tbl-cap: "Shared locations between Ian and Peter"

overlap |>
  select(Location = location,
         `Ian (n)` = ian_n, `Ian (%)` = ian_pct,
         `Peter (n)` = peter_n, `Peter (%)` = peter_pct) |>
  kable(align = c("l", "r", "r", "r", "r"))
```

\vspace{0.2cm}

- `{r} round(overlap_ian / 232 * 100, 1)`% of Ian's customers and `{r} round(overlap_peter / 209 * 100, 1)`% of Peter's customers fall in shared locations.
- **However, actual competitive overlap is minimal** — in each shared location, one rep dominates while the other has only 1–3 customers.
- No immediate need to reassign territories based on overlap alone.

# Risk assessment & recommendations

## Concentration risk summary

\begin{center}
\begin{tabular}{lcc}
\toprule
\textbf{Risk factor} & \textbf{Severity} & \textbf{Affected customers} \\
\midrule
Peter's Ibadan/Oyo concentration & \textcolor{red}{\textbf{HIGH}} & 115 \\
Company-wide Ibadan/Oyo dependence & \textcolor{red}{\textbf{HIGH}} & 711 \\
Ian's broad but thin coverage & \textcolor{orange}{\textbf{MEDIUM}} & 232 \\
Customer Office unassigned accounts & \textcolor{orange}{\textbf{MEDIUM}} & 341 \\
Data quality gaps (addresses, phones) & \textcolor{orange}{\textbf{MEDIUM}} & 247+ records \\
Bottom 11 reps with <30 customers & \textcolor{yellow!80!black}{\textbf{LOW}} & <330 combined \\
\bottomrule
\end{tabular}
\end{center}

## Recommendations for leadership

**Immediate actions (0–3 months):**

1. Assign a backup rep for **Peter's Ibadan/Oyo territory** to mitigate single-person dependency.
2. Audit the **341 Customer Office accounts** — redistribute to named sales reps.
3. Launch a **data cleanup initiative** to fill missing addresses and phone numbers.

\vspace{0.2cm}

**Medium-term actions (3–6 months):**

4. Develop a **succession plan for Ian** — his 31-location coverage is a key-person risk.
5. **Rebalance workloads** — top reps are overloaded while others are underutilised.
6. Evaluate underperforming reps (fewer than 30 customers) for coaching or reassignment.

\vspace{0.2cm}

**Strategic actions (6–12 months):**

7. Explore **geographic expansion** into under-served northern regions.
8. Implement a **CRM system** for better territory management and reporting.

## Thank you

\begin{center}
\vspace{1cm}
{\Large\textbf{Customer \& Sales Rep Analysis}}

\vspace{0.5cm}
Prepared by: \textbf{Akinbami Oluwashola}

For: \textbf{Agro Bar Magen Nigeria LTD}

\vspace{0.5cm}
June 2026

\vspace{1cm}
{\small Questions or feedback? Please reach out to the analyst.}
\end{center}
